%********** Preface **********

\chapter{Preface to the Second Edition}


In the thirty-nine years since Mary Garnett published Bright Leaf, ??.  Its descriptions of tobacco cultivation have been cited in two works of academic history.  Most important of all, it has been passed around and treasured as a family heirloom.

During that time, however, the number of Henry and Julia's descendants has grown far beyond the number of extant copies of Bright Leaf.  The descendant chart has grown incomplete with new generations, and while many copies have been updated with pencil markings...  New sources on Henry and Julia's lives and home have surfaced and the advent of the Internet has X genealogical research

Our approach has been to preserve Mary Garnett's text of Bright Leaf, while updating and extending it through new appendices.  A few typographic errors in the original have been silently corrected, though we fear we may have silently introduced others.  The demands of typesetting have required repositioning some of the charts and illustrations within the text.  Probably the most extensive change has been to the family chart, as the format in the original was not capable of supporting additional generations while remaining readable.

The editors would like to record their thanks to Ellen Wyttenbach for the permission to reprint her mother's work, to Dale Phillips for her leg-work tracking down family members, and to Alison Williams both for her enthuastic support and for coming up with the idea in the first place.

Ben Brumfield and Barbara Lawrence


\pagebreak
